---
title: "Apply"
author: "Vincent Schroeder"
date: "2024-11-21"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. apply()
Purpose: Used to apply a function to the rows or columns of a matrix or 2D array.

Syntax: apply(X, MARGIN, FUN, ...)

X: The matrix or array.
MARGIN: 1 for rows, 2 for columns.
FUN: The function to apply.
Example: Calculate the row sums of a matrix.


```{r}
mat <- matrix(1:9, nrow = 3)
print(mat)
paste("rowsum:", apply(mat, 1, sum))  # Sum across rows
paste("colsums:", apply(mat, 2, sum))  # Sum across rows
```

# 2. lapply()
Purpose: Applies a function to each element of a list and returns a list.

Syntax: lapply(X, FUN, ...)

X: A list or vector.
FUN: The function to apply.
Example: Square each element in a list.

```{r}
lst <- list(1, 2, 3)
print(lapply(lst, function(x) x^2))
print(unlist(lapply(lst, function(x) x^2)))

```

# 3. sapply()
Purpose: Simplified version of lapply() that tries to simplify the result to a vector or matrix, if possible.

Syntax: sapply(X, FUN, ..., simplify = TRUE)

X: A list or vector.
FUN: The function to apply.
Example: Get the length of each element in a list.

```{r}
lst <- list(c(1, 2), c(3, 4, 5), 6, 3:55)
print(sapply(lst, length))

vct = c(list(1:4), list(3:77))
print(sapply(vct, length))
```
# 4. vapply()
Purpose: Similar to sapply(), but it allows the user to predefine the return type, making it safer by preventing unintended simplification.

Syntax: vapply(X, FUN, FUN.VALUE, ..., USE.NAMES = TRUE)

X: A list or vector.
FUN: The function to apply.
FUN.VALUE: A template for the output type.
Example: Get the length of each element in a list, returning a vector of a specific type.

```{r}
lst <- list(c(1, 2), c(3, 4, 5), 6)
vapply(lst, length, FUN.VALUE = integer(1))

```

# 5. mapply()
Purpose: Multivariate version of sapply(). It applies a function to the corresponding elements of multiple lists or vectors.

Syntax: mapply(FUN, ..., MoreArgs = NULL)

FUN: The function to apply.
...: Lists or vectors.
MoreArgs: Additional arguments passed to the function.
Example: Add corresponding elements of two vectors.

```{r}
vec1 <- c(1, 2, 3)
vec2 <- c(4, 5, 6)
mapply(sum, vec1, vec2)

```

# 6. tapply()
Purpose: Applies a function over subsets of a vector, grouped by another vector (factor or categorical data).

Syntax: tapply(X, INDEX, FUN, ...)

X: The vector.
INDEX: The grouping factor or categorical variable.
FUN: The function to apply.
Example: Calculate the mean of a numeric vector grouped by a factor.

```{r}
values <- c(1, 2, 3, 4, 5, 6)
groups <- factor(c('A', 'A', 'B', 'B', 'B', 'A'))
tapply(values, groups, mean)

```
# 7. rapply()
Purpose: Recursive version of lapply(), applying a function to each element of a list, including elements of nested lists.

Syntax: rapply(X, FUN, classes = "ANY", how = "replace", ...)

X: The list or nested list.
FUN: The function to apply.
classes: Specify the class to apply the function to.
how: Whether to "replace" or "list" the results.
Example: Apply the sqrt() function to all numeric elements in a nested list.

```{r}
nested_lst <- list(a = 1:3, b = list(c = 4:6, d = 7))
rapply(nested_lst, sqrt, classes = "numeric", how = "replace")

```

# 8. eapply()
Purpose: Applies a function to each element in an environment.

Syntax: eapply(env, FUN, ...)

env: The environment.
FUN: The function to apply.
Example: List all objects in the environment and their lengths.

```{r}
e <- new.env()
e$a <- 1:3
e$b <- 4:63
e$c = c(1, 2, 3, NA, 4, 5)
eapply(e, mean, na.rm = T)

```


# 9 typeapply()

Purpose: Applies a function to each element in an list with a specified type

Syntax: typeapply(li, fun, type, ...)

li: The list to apply the function on each element. Can be a dataframe to perform column wise operations
FUN: The function to apply.
...: Further arguments for the function
Example: Mean standardise all columns in the iris dataframe, ignoring NAs


```{r}
typeapply = function(li, fun, type, ...) {
  for (i in seq_along(li))
    if (typeof(li[[i]]) == type)
      li[[i]] = fun(li[[i]], ...)
    li
}

df = iris
df[1,1] = NA
typeapply(df, function(x) x - mean(x, na.rm = T), "double")
#typeapply(df, function(x) x - mean(x), "double")
```

